---
title: "genericML_md"
author: "Brandon Williams"
date: "2024-10-31"
output: html_document
---

First, let's load in the data and define some global variables to be used throughout.

```{r setup, include=FALSE}

### Data, global variables, and packages

knitr::opts_chunk$set(echo = TRUE)

# load data
studysample <- read.csv(here("data/studysampleclean.csv"))

# Define global variable list
varlist <- c("presence", 
             "evmar05v3", 
             "evpreg05v3", 
             "evpregunmar05v3", 
             "evunpregmar05v3", 
             "dropout05v3",
             "dropout07v2", 
             "evmar07v2", 
             "evpreg07v2", 
             "evpregunmar07v2", 
             "evunpregmar07v2"
             )
treatlist <- c("Uonly", 
               "Honly",
               "UH"
               )
controllist <- c("yrbirth_all", 
                 "yrbirth_missing", 
                 "date05v3", 
                 "date07v2", 
                 "schsize", 
                 "stratum")

# newest version of GenericML not yet on CRAN so install from here (instead of include.R)
# devtools::install_github("mwelz/GenericML")
library("GenericML")

```

Now let's explore some of the data and determine what we might be missing. 

```{r}

# Replace blanks with NA
studysample[studysample == ""] <- NA

# Count the number of NAs per column
na_counts <- sapply(studysample, function(x) sum(is.na(x)))

# Sort na_counts in ascending order and display the result
na_counts_sorted <- sort(na_counts)
print(na_counts_sorted)

# it appears that there were roughly 7000 people ommitted from the survey in general. So let's keep any NAs under 7000 count


# keep only if less than 2000 NAs
na_counts <- na_counts[na_counts < 7000]



```

Most of these are outcome variables or survey questions taken after treatment. If might be possible to keep survey questions that have stability as relevant, e.g. house characteristics or ROSCA. Let's add some school variables for additional heterogeneity:

```{r}
school <- read_dta(here("data/school_info.dta"))
studysample <- merge(studysample, school, 
                     by = "schoolid",
                     suffixes = c("", ".y")
                     )

controllist <- c("yrbirth_all", 
                 "yrbirth_missing", 
                 "date05v3", 
                 "date07v2", 
                 "schsize", 
                 "stratum",
                 "meanexp",                # teacher experience
                 "latrine_2004",           # number of latrines in the school
                 "sdkcpe",                 # school score ranking in SDs
                 "ratio02"                 # school gender ratio
                 )

```


``` {r}
### Pre-process data

# create a control (non-treated units) vector

studysample <- studysample %>% 
  mutate(control = case_when(
    Utreat.y == 0 & HIVtreat.y == 0 ~ 1,
    TRUE ~ 0
  ))

# reduce sample to necessary vectors

sampleselect <- studysample %>%
  select(all_of(varlist),                   # select relevant columns
         all_of(controllist), 
         all_of(treatlist),
         schoolid
         ) %>% 
  filter(complete.cases(.))                 # remove NA

# create treatment vectors

Uonly <- sampleselect$Uonly                                       
Honly <- sampleselect$Honly
UH <- sampleselect$UH

# Remove columns that correspond to treatment and save as matrix

samplematrix <- data.matrix(sampleselect[, !(names(sampleselect) %in% c(varlist, treatlist))])

# assign ordering to birth year

yrbirth_all <- as.numeric(factor(sampleselect$yrbirth_all))

```


```{r}

### ML Parameters 
 
# specify learners

learners <-
  c("random_forest",                                              # Random Forest  
    "mlr3::lrn('cv_glmnet', s = 'lambda.min', alpha = 0.5)",      # Elastic Net Regularization on generalized linear model
    "mlr3::lrn('svm')",                                           # Support vector machine
    "mlr3::lrn('xgboost')")                                       # Gradient Boost

# Parameterize
seed <- 1833
cutoffs <- c(0.2, 0.4, 0.6, 0.8)
X1 <- setup_X1(funs_Z = c("B", "S")                    # include BCA and CATE controls
               , fixed_effects = yrbirth_all           # cohort effect
               )
vcov <- setup_vcov(
                   # estimator = "vcovCL"              # I was not able to get this to take the vcovCL argument
                   arguments = list(cluster = sampleselect$schoolid)    # cluster on school
                   )        
lps <- c("constant")
num_splits <- 100L
sig_level <- 0.05
parallel <- TRUE
num_cores <- 6L

```


```{r HTE Uonly}

Y <- sampleselect[[varlist[6]]]             # "dropout05v3"
D <- sampleselect[[treatlist[1]]]           # "Uonly"

  
genML <- GenericML(
  Z = samplematrix,                         # matrix
  D = D,                                    # treatment assignment 
  Y = Y,                                    # outcome response variable
  learners_GenericML = learners,            # learners
  learner_propensity_score = lps,           # = 0.5 (RCT)
  num_splits = num_splits,                  # number splits, L is useful for specifying integer and therefore optimizing memory
  quantile_cutoffs = cutoffs,               # grouping
  significance_level = sig_level,           # significance level
  X1_BLP = X1, X1_GATES = X1,               # regression setup
  vcov_BLP = vcov, vcov_GATES = vcov,       # covariance setup
  parallel = parallel,                      # parallelization
  # num_cores = num_cores, 
  seed = seed                               # RNG seed
)

```
Were we able to fix the propensity score error before? Suddenly I don't remember if we solved it or just figured out how it was coming up. 

```{r HTE Uonly}
# BLP
results_BLP <- get_BLP(genML, plot = TRUE)
results_BLP       # print method
plot(results_BLP) # plot method


```

```{r HTE}
# GATES
results_GATES <- get_GATES(genML, plot = TRUE)
results_GATES
plot(results_GATES)


```

``` {r HTE}

# CLAN
results_CLAN <- get_CLAN(genML, variable = "schsize", plot = TRUE)
results_CLAN
plot(results_CLAN)
```

```{r HTE}
# best learners
get_best(genML)
```
Since xgboost is giving us the best results, let's specify that as the only learner and speed up computation time for a loop.

```{r Uonly loop}
learners <-
  c("mlr3::lrn('xgboost')")                                       # Gradient Boost

# Start timer
start_time <- Sys.time()

# Initialize a list to store results
genML_results <- list()

# Loop over the first 11 entries in varlist
for (i in 1:11) {
  # Select outcome variable Y and treatment assignment D
  Y <- sampleselect[[varlist[i]]]
  D <- sampleselect[[treatlist[1]]]  # Keeping constant across iterations

  # Run the GenericML function for each Y
  genML <- GenericML(
    Z = samplematrix,                         # matrix
    D = D,                                    # treatment assignment 
    Y = Y,                                    # outcome response variable
    learners_GenericML = learners,            # learners
    learner_propensity_score = lps,           # = 0.5 (RCT)
    num_splits = num_splits,                  # number splits, L is useful for specifying integer and therefore optimizing memory
    quantile_cutoffs = cutoffs,               # grouping
    significance_level = sig_level,           # significance level
    X1_BLP = X1, X1_GATES = X1,               # regression setup
    vcov_BLP = vcov, vcov_GATES = vcov,       # covariance setup
    parallel = parallel,                      # parallelization
    # num_cores = num_cores, 
    seed = seed                               # RNG seed
  )
  
  # Save the result for each iteration in the list
  genML_results[[i]] <- genML

  # Display progress message
  cat("Completed iteration:", i, "of 11\n")
}

# End timer and calculate elapsed time
end_time <- Sys.time()
total_time <- end_time - start_time
cat("Total time taken:", total_time, "\n")


```

```{r HTE Uonly}
# BLP

# Loop over each genML result stored in genML_results
for (i in 1:11) {
  # Get BLP results with plot enabled
  results_BLP <- get_BLP(genML_results[[i]], plot = TRUE)
  
  # Plot the results
  print(plot(results_BLP))
}


```


```{r Uonly}
# GATES

# Loop over each genML result stored in genML_results
for (i in 1:11) {
  # Get BLP results with plot enabled
  results_GATES <- get_GATES(genML_results[[i]], plot = TRUE)
  
  # Plot the results
  print(plot(results_GATES))
}

```
```{r }
learners <-
  c("mlr3::lrn('xgboost')")                                       # Gradient Boost

# Start timer
start_time <- Sys.time()

# Initialize a list to store results
genML_results_Honly <- list()

# Loop over the first 11 entries in varlist
for (i in 1:11) {
  # Select outcome variable Y and treatment assignment D
  Y <- sampleselect[[varlist[i]]]
  D <- sampleselect[[treatlist[2]]]  # Honly

  # Run the GenericML function for each Y
  genML <- GenericML(
    Z = samplematrix,                         # matrix
    D = D,                                    # treatment assignment 
    Y = Y,                                    # outcome response variable
    learners_GenericML = learners,            # learners
    learner_propensity_score = lps,           # = 0.5 (RCT)
    num_splits = num_splits,                  # number splits, L is useful for specifying integer and therefore optimizing memory
    quantile_cutoffs = cutoffs,               # grouping
    significance_level = sig_level,           # significance level
    X1_BLP = X1, X1_GATES = X1,               # regression setup
    vcov_BLP = vcov, vcov_GATES = vcov,       # covariance setup
    parallel = parallel,                      # parallelization
    # num_cores = num_cores, 
    seed = seed                               # RNG seed
  )
  
  # Save the result for each iteration in the list
  genML_results_Honly[[i]] <- genML

  # Display progress message
  cat("Completed iteration:", i, "of 11\n")
}

# End timer and calculate elapsed time
end_time <- Sys.time()
total_time <- end_time - start_time
cat("Total time taken:", total_time, "\n")


```
```{r HTE Honly}
# BLP

# Loop over each genML result stored in genML_results
for (i in 1:11) {
  # Get BLP results with plot enabled
  results_BLP <- get_BLP(genML_results[[i]], plot = TRUE)
  
  # Plot the results
  print(plot(results_BLP))
}


```

```{r HTE}
# GATES

# Loop over each genML result stored in genML_results
for (i in 1:11) {
  # Get BLP results with plot enabled
  results_GATES <- get_GATES(genML_results[[i]], plot = TRUE)
  
  # Plot the results
  print(plot(results_GATES))
}

```
Next steps:
1. Align results based on inputs
2. CLAN breakdown of het treat effects if necessary
3. (Partially done) Additional covariates: any more reasonable characteristics we could find?
4. (Done) GenML on other outcome/treatment variables